---
title: "The clifford package: Clifford algebra in R"
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![Build Status](https://travis-ci.org/RobinHankin/clifford.svg?branch=master)](https://travis-ci.org/RobinHankin/clifford)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/clifford)](https://cran.r-project.org/package=clifford)
[![Rdoc](https://www.rdocumentation.org/badges/version/clifford)](https://www.rdocumentation.org/packages/clifford)
[![Codecov test coverage](https://codecov.io/gh/RobinHankin/clifford/branch/master/graph/badge.svg)](https://codecov.io/gh/RobinHankin/clifford/branch/master)
<!-- badges: end -->

# Overview

The `clifford` package provides R-centric functionality for working
with Clifford algebras of arbitrary dimension and signature.  A
detailed vignette is provided in the package.

# Installation

You can install the released version of the clifford package from
[CRAN](https://CRAN.R-project.org) with:

```{r, message=FALSE}
# install.packages("clifford")  # uncomment this to install the package
library("clifford")
set.seed(0)
```

# The `clifford` package in use

The basic creation function is `clifford()`, which takes a list of
basis blades and a vector of coefficients:

```{r cliff1}
(a <- clifford(list(1,2,1:4,2:3),1:4))
(b <- clifford(list(2,2:3,1:2),c(-2,3,-3)))
```

So `a` and `b` are multivectors.  Clifford objects are a vector space
and we can add them using `+`:

```{r,cliff2}
a+b
```

See how the `e2` term vanishes and the `e_23` term is summed.
The package includes a large number of products:


```{r,cliffprods}
a*b        # geometric product (also "a % % b")
a %^% b    # outer product
a %.% b    # inner product
a %star% b # scalar product
a %euc% b  # Euclidean product
```

The package can deal with non positive-definite inner products.
Suppose we wish to deal with an inner product of 

\[
\begin{pmatrix}
+1 & 0 & 0 & 0 & 0\\
 0 &+1 & 0 & 0 & 0\\
 0 & 0 &+1 & 0 & 0\\
 0 & 0 & 0 &-1 & 0\\
 0 & 0 & 0 & 0 &-1
\end{pmatrix}
\]

where the diagonal is a number of $+1$ terms followed by a number of
$-1$ terms.  The package idiom for this would be to use `signature()`:

```{r,clifsig}
signature(3)
```

Function `signature()` is based on `lorentz::sol()` and its argument
specifes the number of basis blades that square to $+1$, the others
squaring to $-1$.  Thus $e_1^2=e_2^2=e_3^2=1$ and $e_4^2=e_5^2=-1$:

```{r,showsig}
basis(1)
basis(1)^2
basis(4)
basis(4)^2
```

The package uses the STL map class with dynamic bitset keys for
efficiency and speed and can deal with objects of arbitrary
dimensions.  Thus:

```{r}
options("basissep" = ",")
(x <- rcliff(d=20))
summary(x^3)
```


# References

  - D. Hestenes 1987. _Clifford algebra to geometric calculus_, Kluwer.
  - J. Snygg 2010. _A new approach to differential geometry using Clifford's geometric algebra_.  Berghauser.
  - C. Perwass 2009. _Geometric algebra with applications
      in engineering_. Springer.

# Further information

For more detail, see the package vignette

```vignette("clifford")```
