---
title: "The clifford package: Clifford algebra in R"
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![Build Status](https://travis-ci.org/RobinHankin/clifford.svg?branch=master)](https://travis-ci.org/RobinHankin/clifford)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/clifford)](https://cran.r-project.org/package=clifford)
[![Rdoc](http://www.rdocumentation.org/badges/version/clifford)](http://www.rdocumentation.org/packages/clifford)
[![Codecov test coverage](https://codecov.io/gh/RobinHankin/clifford/branch/master/graph/badge.svg)](https://codecov.io/gh/RobinHankin/clifford/branch/master)
<!-- badges: end -->

# Overview

The ``clifford` package provides R-centric functionality for working
with Clifford algebras of arbitrary dimension and signature.  A
detailed vignette is provided in the package.

# Installation

You can install the released version of the clifford package from
[CRAN](https://CRAN.R-project.org) with:

```{r, message=FALSE}
# install.packages("clifford")  # uncomment this to install the package
library("clifford")
set.seed(0)
```

# The `clifford` package in use

```{r cliff1}
a <- rcliff()
a
```

# References

  - D. Hestenes 1987. _Clifford algebra to geometric calculus_, Kluwer.
  - J. Snygg 2010. _A new approach to differential geometry using Clifford's geometric algebra_.  Berghauser.

# Further information

For more detail, see the package vignette

```vignette("clifford")```
