---
title: "Clifford involutions following Hitzer and Sangwine"
author: "Robin K. S. Hankin"
date: "02/09/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
bibliography: clifford.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{involutions}
  %\usepackage[utf8]{inputenc}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clifford inverses

Hitzer and Sangwine set up a number of involutions which I reproduce
for convenience below.  Given $M\in Cl(p,q)$ and 

\newcommand{\abr}[1]{\left\langle #1\right\rangle}

$$M=
\abr{M}_0+
\abr{M}_1+
\abr{M}_2+\cdots+\abr{M}_0
$$

we have

* The main grade involution
$\displaystyle\widehat{M}=\sum_{k=0}^n(-1)^k\abr{M}_k\qquad\mbox{\tt gradeinv(M)}$
* Reversion $\displaystyle\widetilde{M}=\sum_{k=0}^n(-1)^{k(k-1)/2}\abr{M}_k\qquad\mbox{\tt rev(M)}$
* Clifford conjugation $\displaystyle\overline{M}=\sum_{k=0}^n(-1)^{k(k+1)/2}\abr{M}_k\qquad\mbox{\tt cliffconj(M)}$
* Grade specific maps $\displaystyle m_{\overline{j},\overline{k}}(M)=M-2\left(\abr(M)_j+\abr{M}_k\right)\qquad\mbox{\tt neg(M)}$


H\&S  assert that

$$\overline{M} = \widehat{\widetilde{M}} = \widetilde{\widehat{M}}
=
\sum_{k=0}^n(-1)^{k(k+1)/2}\abr{M}_k
$$

which we may verify numerically:

```{r,loadcliffordlibrary,echo=TRUE,print=FALSE,message=FALSE}
library(clifford)
```


```{r,label=equation2.6}
(M <- rcliff())
a1 <- cliffconj(M)
a2 <- gradeinv(rev(M))
a3 <- rev(gradeinv(M))
is.zero(a2-a1) & is.zero(a3-a1)
```

We now consider the case $p+q=3$.  If $x\in Cl(p,q)$ with $p+q=3$ then
equation 6.2 asserts that $x\overline{x}=r_0 +r_3 e_1e_2e_3$ for some
$r_0,r_3\in\mathbb{R}$:

```{r,label=equation6.2}
(x <- rcliff(d=3,g=3))
x*cliffconj(x)
```

and equation 6.3 asserts that $x\overline{x}(x\overline{x})^\sim\in\mathbb{R}$:

```{r,label=equation6.3}
	f <- function(x){
	jj <- x*cliffconj(x)
	is.real(jj*rev(jj))
}
```

```{r}
signature(0)
f(rcliff(d=3,g=3))
signature(1)
f(rcliff(d=3,g=3))
signature(2)
f(rcliff(d=3,g=3))
signature(3)
f(rcliff(d=3,g=3))
```
