\documentclass{article}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\newcommand{\cliff}[1]{\ensuremath{C\ell\left(#1\right)}}
\newcommand{\ei}[1]{\ensuremath{{\bf e}_{#1}}}
\begin{document}

Manuscript Number: AACA-24-00133

\section*{Rebuttal to reviewers' comments}

Below, I reproduce the comments made to me on 23 March 2025.  I have
retrofitted the email to \LaTeX, and made minor changes to the
typesetting, but have not changed the sense of the comments.  Replies
to the issues, and my own observations and general comments, are in
\textcolor{blue}{blue}.  I have indicated changes to the manuscript
where appropriate.\\

\vspace*{1cm}

\noindent\textcolor{blue}{{\bf Executive Summary.\ \ } All the
  comments are perfectly reasonable and I have implemented them to the
  best of my ability.  I believe the manuscript is much improved as a
  result, and recommend it to you.}
\vspace*{1cm}

\begin{itemize}
\item give more details about the data structure and products.  For
  example, how the author's data structure handle $k$-vectors,
  $k$-blades, general multivectors with a focus on computational and
  memory complexity.  This is important to know for the GA
  implementation community.

\textcolor{blue}{Discussion added at the end of section 2.  In short,
  insertion, deletion, and lookup are all $\mathcal{O}(\log n)$
  operations and I interpret this for $k$-vectors as suggested by the
  review.  I also give some $\mathcal{O}()$ results for geometric and
  outer products as suggested.}

\item give a comparison with state-of-the-art methods.  The author's
  package is the only one written in R but it is not the only GA
  implementation and it is not the only symbolic computation
  package.  For example, other methods use the bitset representation of
  blades (like Gaigen), this would be good to recall this in the
  paper.  This comparison would provide valuable context and highlight
  the package's contributions.

\textcolor{blue}{Computational implementations of Clifford algebra are
  legion.  I think the reviewer is directing me to interpret "symbolic
  computation package" very much more broadly than I did in earlier
  drafts.  This is fine.  I have added a short discussion to some of
  the more germane implementations, including the three citations
  [a-c] that the review suggests.}
  
\item add some experimental results on how to use the package e.g. in
  computer graphics (discrete differential geometry) or physics or
  statistics.  In the following paragraphs, I explain the above points
  in more details

\textcolor{blue}{Brief section on conformal geometry added, showing
  how to use the package to identify the sphere that passes through
  four given points}

\item Section 2 should contain more details about the data structure
  used: the potential problems of using an associative array (time
  access).

\item Appendix B should not be in the appendix.  It is a valuable
  addition to the Section 2.

  \textcolor{blue}{Done}
  
\item The presented state-of-art methods should include other symbolic
  implementation like Clifford implementation (Python), Galop..., see
  [a,b].

\item In section 2, "The package uses standard disordR discipline,
  which efficiently exploits the fact that coefficients are stored as
  an unordered map.  Full details are given in [6]".  The unordered
  map data structure should be described, the reader should not have
  to look at [6] to understand the authors multivector data
  structure. Please complete this section.

\textcolor{blue}{Done (although it made more sense to include it as an
  appendix as the requisite concepts have to be developed first).}
  
\item Since the used data structure is an unordered map, the
  computational cost of finding all the blades of a given grade is not
  a $\mathcal{O}(\log(n))$ operation.  The author should add more
  details about the time complexity of operations like computing the
  outer products of a $k$-vector and a $k'$-vector.  These kind of
  operations are often used in Conformal Geometric Algebra.

\item About Section 8, the author could give an example of an
  application in either differential geometry/CG/CV and how it is
  written with clifford, see [c,d].

\end{itemize}

\section*{References}

\begin{description}
\item{[a]} \url{https://github.com/ga/awesome-geometric-algebra}
\item{[b]} Hildenbrand, D., Pitt, J., \& Koch, A. (2010).
  Gaalop---high performance parallel computing based on conformal
  geometric algebra.  Geometric Algebra Computing: in Engineering and
  Computer Science, 477-494.
\item{[c]} Breuils, S., Nozick, V., Sugimoto, A., \& Hitzer,
  E. (2018).  Quadric conformal geometric algebra of
  $\mathbb{R}^{9,6}$.  Advances in Applied Clifford Algebras, 28,
  1-16.
\item{[d]} Hildenbrand, D. (2012, September).  Foundations of geometric
  algebra computing.  In AIP Conference Proceedings (Vol. 1479, No. 1,
  pp. 27-30).  American Institute of Physics.
\end{description}

\end{document}
